---
title: "ISYE 6402 Project EDA and ARIMA"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
```{r setup, include = FALSE}

# Set up the default parameters
# 1. The code block will be shown in the document
# 2. set up figure display size
# 3. turn off all the warnings and messages

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 4)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

## Background



##Data import and cleaning

```{r}
## Libraries used within this homework are uploaded here
library(zoo,warn.conflicts=FALSE)
library(lubridate,warn.conflicts=FALSE)
library(mgcv,warn.conflicts=FALSE)
library(rugarch,warn.conflicts=FALSE)
library(readxl)
library(xts)
```

```{r}
#importing the data
data_small <- read_excel("Data_Template.xlsx")
data_full <- read.csv("final.csv", head = TRUE)

data <- data_small

#cleaning the data

#dates to date format
data$Date<-as.Date(data$Date, format='%Y-%m-%d/')
data = data[order(data$Date), ]

?xts()
#xts format
data$SPY <- xts(data$SPY, data$Date)
data$DJ <- xts(data$DJ, data$Date)


plot(data$SPY, main="S&P 500", ylab="Index")
plot(data$DJ, main="Dow Jones", ylab=" Index")
 
```

## Exploratory Data Analysis and Trend Estimation

```{r}
# time point variables
time.pts = c(1:length(data$SPY))
time.pts = c(time.pts - min(time.pts))/max(time.pts)
start_date = data$Date[order(data$Date)][1]
start_date_c = c(as.numeric(format(start_date, format="%Y")), 
                 as.numeric(format(start_date, format="%m")),
                 as.numeric(format(start_date, format="%d")))

# Spline model
gam.fit.SP = gam(data$SPY~s(time.pts))
SP.fitted = ts(fitted(gam.fit.SP),start=start_date_c,frequency=12)

gam.fit.DJ = gam(data$DJ~s(time.pts))
DJ.fitted = ts(fitted(gam.fit.DJ),start=start_date_c,frequency=12)

# plotting

ts.plot(ts(data$SPY, start=start_date_c,frequency=12) ,ylab="S&P 500", main = "S&P Trend Prediction")
lines(SP.fitted,lwd=2,col="red", type='l')
grid()

ts.plot(ts(data$DJ, start=start_date_c,frequency=12),ylab="Dow Jones", main = "Dow Jones Trend Prediction")
lines(DJ.fitted,lwd=2,col="red", type='l')
grid()

```


The trends and fluctuations of Dow Jones and S&P are largely the same. There is no observable seasonality. 

# Differenced data
```{r}
start_date =  data$Date[order(data$Date)][2]
start_date_c = c(as.numeric(format(start_date, format="%Y")), 
                 as.numeric(format(start_date, format="%m")),
                 as.numeric(format(start_date, format="%d")))

diff.SP = ts(diff(data$SPY)[-1], start=start_date_c,frequency=12)
diff.DJ =  ts(diff(data$DJ)[-1], start=start_date_c,frequency=12)

ts.plot(diff.DJ, col="red", main="Compare differenced data of S&P and DowJones")
lines(diff.SP, col="blue")
grid()

ts.plot(diff.SP, col="blue", main="Differenced S&P")
ts.plot(diff.DJ, col="red", main="Differenced DowJones")


```
## ARIMA Model

```{r}

n_forward = 2                           ###period forward want to predict
n = length(data$SPY)
nfit = n-n_forward                    ###number rows fit to training set
SPY.train = data$SPY[1:nfit]
SPY.test = data$SPY[(nfit+1):n]

DJ.train = data$DJ[1:nfit]
DJ.test = data$DJ[(nfit+1):n]

test_modelA <- function(p, d, q, data) {
  mod = arima(data, order = c(p, d, q), method = "ML")
  current.aic = AIC(mod)
  df = data.frame(p, d, q, current.aic)
  names(df) <- c("p","d","q","AIC")
  print(paste(p,d,q,current.aic,sep=" "))
  return(df)
}


##### S & P Model #####
orders.SP = data.frame(Inf,Inf,Inf,Inf)
names(orders) <- c("p","d","q","AIC")

for (p in 0:8) {
  for (d in 0:2) {
    for (q in 0:8) {
      possibleError <- tryCatch(
        orders.SP <- rbind(orders, test_modelA(p, d, q, data=SPY.train)),
        error = function(e) {e}
      )
      if (inherits(possibleError, "error"))
        next
    }
  }
}
orders.SP <- orders.SP[order(-orders.SP$AIC), ]
tail(orders.SP)



##### Dow Jones #####
orders.DJ = data.frame(Inf,Inf,Inf,Inf)
names(orders) <- c("p","d","q","AIC")

for (p in 0:8) {
  for (d in 0:2) {
    for (q in 0:8) {
      possibleError <- tryCatch(
        orders.DJ <- rbind(orders, test_modelA(p, d, q, data=SPY.train)),
        error = function(e) {e}
      )
      if (inherits(possibleError, "error"))
        next
    }
  }
}
orders.DJ <- orders.DJ[order(-orders.DJ$AIC), ]
tail(orders.DJ)



```


# Residual Analysis of ARIMA Model
```{r}
## Final Model
mod.SP.1 = arima(SPY.train,order = c(7, 2 ,8), method='ML')
# mod.SP.2 = arima(SPY.train,order = c(4, 2 ,5), method='ML')

mod.DJ.1 = arima(DJ.train,order = c(7, 2 ,8), method='ML')
# mod.DJ.2 = arima(DJ.train,order = c(4, 2 ,5), method='ML')

mod.SP.1
mod.DJ.1


### Residual Analysis
resid.SP <- resid(mod.SP.1)
resid.DJ <- resid(mod.DJ.1)

### SP

plot(resid.SP, ylab='Standardized Residuals',type='o',main="Residual Plot" )
abline(h=0, col='red')
acf(as.vector(resid.SP), lag.max=12*5, main="ACF: S&P ARIMA Residual")
pacf(as.vector(resid.SP), lag.max=12*5, main="PACF: S&P ARIMA Residual")
hist(resid.SP,xlab='Standardized Residuals',main='Histogram: Residuals')
qqnorm(resid.SP)
qqline(resid.SP)


### DJ

plot(resid.DJ, ylab='Standardized Residuals',type='o',main="Residual Plot" )
abline(h=0, col='red')
acf(as.vector(resid.DJ), lag.max=12*5, main="ACF: DowJon ARIMA Residual")
pacf(as.vector(resid.DJ), lag.max=12*5, main="PACF: DowJon ARIMA Residual")
hist(resid.DJ,xlab='Standardized Residuals',main='Histogram: Residuals')
qqnorm(resid.DJ)
qqline(resid.DJ)

## hypothesis testing
porder=7
qorder=8
Box.test(mod.SP.1$resid, lag = (porder+qorder+1), type = "Box-Pierce", fitdf = (porder+qorder))
Box.test(mod.SP.1$resid, lag = (porder+qorder+1), type = "Ljung-Box", fitdf = (porder+qorder))

Box.test(mod.DJ.1$resid, lag = (porder+qorder+1), type = "Box-Pierce", fitdf = (porder+qorder))
Box.test(mod.DJ.1$resid, lag = (porder+qorder+1), type = "Ljung-Box", fitdf = (porder+qorder))



```

The p-values of the ljung-box tests are small and reject the null hypothesis at confidence level of 95%. It is plausible that the residuals are correlated for both ARIMA model fitted into S&P and DowJones.

## ARMA-GARCH Model
```{r}

### S&P

# Step 1: Initial ARMA orders
test_modelA1 <- function(p, d, q) {
  mod = arima(diff.SP, order = c(p, d, q), method = "ML") 
  current.aic = AIC(mod)    
  df = data.frame(p, d, q, current.aic)
  names(df) <- c("p", "d", "q", "AIC") 
  print(paste(p,d,q,current.aic,sep=" "))
  return(df)
}

ordersARIMA = data.frame(Inf, Inf, Inf, Inf)
names(ordersARIMA) <- c("p", "d", "q", "AIC")
for (p in 0:8) { 
  for (d in 0:0) {
    for (q in 0:8) { 
      possibleError <- tryCatch(
            ordersARIMA <- rbind(ordersARIMA, test_modelA1(p, d, q)),
            error = function(e) {e}
      )
      if (inherits(possibleError, "error")) 
        next
    } 
  }
}
ordersARIMA <- ordersARIMA[order(-ordersARIMA$AIC), ]
tail(ordersARIMA)

### ARIMA Initial Orders -> (7, 1, 6) is chosen
porder = tail(ordersARIMA, n=1)$p
qorder = tail(ordersARIMA, n=1)$q

# 

# Step 2: Initial GARCH Order
#ARIMA-GARCH: Select GARCH order
test_modelAGG <- function(m,n){
    spec = ugarchspec(variance.model=list(garchOrder=c(m,n)),
                      mean.model=list(armaOrder=c(porder,qorder),
                                      include.mean=T),
                      distribution.model="std")
    fit = ugarchfit(spec, diff.SP, solver = 'hybrid')
    current.bic = infocriteria(fit)[2]
    df = data.frame(m,n,current.bic)
    names(df) <- c("m","n","BIC")
    print(paste(m,n,current.bic,sep=" "))
    return(df)
}

ordersAGG = data.frame(Inf,Inf,Inf)
names(ordersAGG) <- c("m","n","BIC")

for (m in 0:3){
    for (n in 0:3){
        possibleError <- tryCatch(
            ordersAGG<-rbind(ordersAGG,test_modelAGG(m,n)),
            error=function(e) e
        )
        if(inherits(possibleError, "error")) next
    }
}
ordersAGG <- ordersAGG[order(-ordersAGG$BIC),]
tail(ordersAGG)
# 1,1

### GARCH Initial Orders -> (1, 1) is chosen
morder = tail(ordersAGG, n=1)$m
norder = tail(ordersAGG, n=1)$n

#Step 3: ARMA order update
#ARIMA-GARCH: Select ARIMA order
test_modelAGA <- function(p,q){
    spec = ugarchspec(variance.model=list(garchOrder=c(morder,norder)),
                      mean.model=list(armaOrder=c(p,q),
                                      include.mean=T),
                      distribution.model="std")
    fit = ugarchfit(spec, diff.SP, solver = 'hybrid')
    current.bic = infocriteria(fit)[2]
    df = data.frame(p,q,current.bic)
    names(df) <- c("p","q","BIC")
    print(paste(p,q,current.bic,sep=" "))
    return(df)
}

ordersAGA = data.frame(Inf,Inf,Inf)
names(ordersAGA) <- c("p","q","BIC")
for (p in 0:4){
    for (q in 0:4){
        possibleError <- tryCatch(
            ordersAGA<-rbind(ordersAGA,test_modelAGA(p,q)),
            error=function(e) e
        )
        if(inherits(possibleError, "error")) next
    }
}
ordersAGA <- ordersAGA[order(-ordersAGA$BIC),]
tail(ordersAGA)
# Don't choose 0,0 since it's trivial.
#1,0

### ARMA Update -> (0, 1) is chosen since (0, 0) is trivial and not selected. 
#### Also, note that (1, 0) and (0, 1) has the same BIC. 
#### Since we are estimating variance of Z_t in GARCH, 
#### (0, 1) is preferred to (1, 0).
porder = 1; qorder = 0


# Step 4: GARCH order update
test_modelAGG <- function(m,n){
    spec = ugarchspec(variance.model=list(garchOrder=c(m,n)),
                      mean.model=list(armaOrder=c(porder,qorder),
                                      include.mean=T), distribution.model="std")
    fit = ugarchfit(spec, diff.SP, solver = 'hybrid')
    current.bic = infocriteria(fit)[2]
    df = data.frame(m,n,current.bic)
    names(df) <- c("m","n","BIC")
    print(paste(m,n,current.bic,sep=" "))
    return(df)
}

ordersAGG = data.frame(Inf,Inf,Inf)
names(ordersAGG) <- c("m","n","BIC")

for (m in 0:3){
    for (n in 0:3){
        possibleError <- tryCatch(
            ordersAGG<-rbind(ordersAGG,test_modelAGG(m,n)),
            error=function(e) e
        )
        if(inherits(possibleError, "error")) next
    }
}
ordersAGG <- ordersAGG[order(-ordersAGG$BIC),]
tail(ordersAGG)

#### Final ARMA(1, 0)+GARCH(1,1) selected 
morder = tail(ordersAGG, n=1)$m
norder = tail(ordersAGG, n=1)$n

### Final Order Selection
morder = 1
norder = 1
porder = 1
qorder = 0

# Building Final Model
spec.SP = ugarchspec(variance.model=list(garchOrder=c(morder, norder)),
                 mean.model=list(armaOrder=c(porder, qorder),
                 include.mean=T), distribution.model="std")
final.model.SP = ugarchfit(spec.SP, diff.SP, solver = 'hybrid')

## compare Information Criteria
infocriteria(final.model.SP)

## Residual Analysis
resids.final.model = residuals(final.model.SP)
acf(resids.final.model,main="ACF of ARCH Residuals")
acf(resids.final.model^2,main="ACF of Squared ARCH Residuals")
Box.test(resids.final.model,lag=10,type='Ljung')
Box.test(resids.final.model^2,lag=10,type='Ljung')
qqnorm(resids.final.model)
```

```{r}

```

